Communicative pressures at the semantics-pragmatics interface:\\ Learning biases may prevent the lexicalization of pragmatic inferences

### 2016-10-06 paper/paper.tex
references on laboratory experiments in IL and their somewhat choppy inclusion in the main text

The main innovations of the model are its modular separation of expressivity and learnability, allowing for their isolated and combined analysis, as well as the learning process involving a joint inference over types of pragmatic behavior and lexical meaning. The goal to decouple but model both expressivity and learnability has also recently been addressed by \citet{kirby+etal:2015}. In contrast to our proposal, however, Kirby et al. model expressivity as exerting its force only in the production of learning data. This model's expressivity parameter thereby fulfills a similar role  to high values of $\lambda$ and $\alpha$ in making speaker behavior more deterministic. In this way, it ``favors'' unambiguous languages. Crucially, the degree of mutual understanding of interlocutors central to replication and to our notion of expressivity is not taken into consideration. Our reasoning behind its inclusion lies in the empirical and theoretical observation that learnability alone can have the undesirable outcome of favoring functionally defective languages. This outcome is predicted by our analysis and similar effects have also been reported in a number laboratory experiments where the participants' task was to learn and subsequently reproduce the language produced by a previous participant. This process was shown to generally lead to a proliferation of languages that associated a large number of meanings with a single form (\hl{references}). In contrast, adding an experiments involving an interactive component have been found to foster languages that enable interlocutors to distinguish meanings accurately  (\hl{references}; for a review of these studies and further references see \citealt{kirby+etal:2015, tamariz+kirby:2016}). More generally, we think there is an important difference between a language generating learnable linguistic data and its actual performance as a means of information transfer. The former solely depends on the mechanism by which speakers associate form and meaning. The latter additionally hinges on the hearers linguistic experience and her capability to interpret linguistic input based on this experience. In sum, we contend that successful information transfer in a linguistic community is central the adoption of a communication system and that this measure is not faithfully reflected by taking only production into consideration.

>>>>

The main innovations of the model are its modular separation of expressivity and learnability, allowing for their isolated and combined analysis, as well as the learning process involving a joint inference over types of pragmatic behavior and lexical meaning. The goal to decouple but model both expressivity and learnability has also recently been addressed by \citet{kirby+etal:2015}. In contrast to our proposal, however, Kirby et al. model expressivity as exerting its force only in the production of learning data. This model's expressivity parameter thereby fulfills a similar role  to high values of $\lambda$ and $\alpha$ in making speaker behavior more deterministic. In this way, it ``favors'' unambiguous languages. Crucially, the degree of mutual understanding of interlocutors central to replication and to our notion of expressivity is not taken into consideration. That is, while our proposal combines bidirectional horizontal transmission with its vertical and unidirectional counterpart, Kirby et al.'s model only considers the latter's influence. Our reasoning behind the inclusion of the former lies in the empirical and theoretical observation that learnability alone can lead the selection of functionally defective languages, as showcased by the tautological language $L_3$ in our analysis. This outcome has been reported in a number laboratory experiments where the participants' task was to learn and subsequently reproduce the language produced by a previous participant, leading to a proliferation of languages that associated a large number of meanings with a single form (see e.g. \citealt{silvey+etal:2014} and experiment 1 in \citealt{kirby+etal:2008}). In contrast, experiments involving an interactive component have been found to foster languages that enable interlocutors to distinguish meanings accurately  (e.g. \citealt{fay+etal:2013}; for a review of laboratory results under the iterated learning paradigm and further discussion see \citealt{kirby+etal:2015, tamariz+kirby:2016}). It is not evident how to compare these empirical predictions given their variation in meaning spaces, modes of transmission, iterations and feedback given to participants. However, we take these results to suggest that there is an important difference between a language generating learnable linguistic data and its actual performance as a means of information transfer. The former solely depends on the mechanism by which speakers associate form and meaning. The latter additionally hinges on the hearers linguistic experience and her capability to interpret linguistic input based on this experience. In sum, we contend that successful information transfer in a linguistic community is central the adoption of a communication system and that this measure is not faithfully reflected by taking only production into consideration.
